\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-24 paper.qmd
Removed distance diag from Figure 9
Figure 10 initial work

The default number of bins $b=b_1\times b_2$ is computed based on the sample size, by setting $b_1=n^{1/3}$, consistent with the Diaconis-Freedman rule [@freedman1981]. The value of $b_2$ is determined analytically by $b_1, q, r_2$. Values of $b_1$ between $2$ and $b_1 = \sqrt{\frac{n}{2}}$ are allowed. @fig-param-scurve (b) shows the effect of different choices of $b_1$ on the MSE of the fitted model.

@@

@@

@@

benchmark_label_df <- tibble(
  y = benchmark_scurve - 0.03, 
  x = 1.3, 
  label = paste0("benchmark is ", benchmark_scurve))

@@

mse_scurve_b + a1_m_scurve +
  mse_scurve_lwd + distance_hist + 
  plot_layout(ncol=2)
```

>>>>

The default number of bins $b=b_1\times b_2$ is computed based on the sample size, by setting $b_1=n^{1/3}$, consistent with the Diaconis-Freedman rule [@freedman1981]. The value of $b_2$ is determined analytically by $b_1, q, r_2$. Values of $b_1$ between $2$ and $b_1 = \sqrt{\frac{n}{r_2}}$ are allowed. @fig-param-scurve (b) shows the effect of different choices of $b_1$ on the MSE of the fitted model.

@@

mse_scurve_b + a1_m_scurve +
  mse_scurve_lwd + 
  plot_layout(ncol=3)
```

@@

trimesh_scurve <- ggplot() +
  geom_segment(data = tr_from_to_df_scurve,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#33a02c") +
  interior_annotation("a", position = c(0.92, 0.95))

@@

## Benchmark 1

distance_df_small_edges_scurve1 <- distance_scurve |>
  filter(distance < benchmark_scurve)

tr_from_to_df_scurve1 <- inner_join(
  tr_from_to_df_scurve, distance_df_small_edges_scurve1,
  by = c("from", "to"))

trimesh_scurve_removed1 <- ggplot() +
  geom_segment(data = tr_from_to_df_scurve1,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#33a02c") +
  interior_annotation("c", position = c(0.92, 0.95))

## Benchmark 2

benchmark_scurve2 <- 0.6

distance_df_small_edges_scurve2 <- distance_scurve |>
  filter(distance < benchmark_scurve2)

tr_from_to_df_scurve2 <- inner_join(
  tr_from_to_df_scurve, distance_df_small_edges_scurve2,
  by = c("from", "to"))

trimesh_scurve_removed2 <- ggplot() +
  geom_segment(data = tr_from_to_df_scurve2,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#33a02c") +
  interior_annotation("b", position = c(0.92, 0.95))

@@

```{r}
#| fig-width: 6
#| fig-height: 4
#| fig-cap: "long edge removal"

trimesh_scurve + trimesh_scurve_removed2 + trimesh_scurve_removed1 +
  plot_layout(ncol = 3)
```

### 2024-09-24 paper.tex
Updated the Figure 15 title
Added proj for twisted patterns
Removed distance diag from Figure 9
Figure 10 initial work

The default number of bins \(b=b_1\times b_2\) is computed based on the
sample size, by setting \(b_1=n^{1/3}\), consistent with the
Diaconis-Freedman rule \citep{freedman1981}. The value of \(b_2\) is
determined analytically by \(b_1, q, r_2\). Values of \(b_1\) between
\(2\) and \(b_1 = \sqrt{\frac{n}{2}}\) are allowed.
Figure~\ref{fig-param-scurve} (b) shows the effect of different choices
of \(b_1\) on the MSE of the fitted model.

@@

@@

@@

\caption{\label{fig-mnist-mse}Assessing which of the 5 NLDR layouts on
the MNIST digit 1 data is the better representation using MSE for
varying binwidth (\(a_1\)). Colour used for the lines and points in the
left plot and in the scatterplots represents NLDR layout (a-e). Layout c
is universally poor. Layouts a that show two close clusters are
universally suboptimal. Layout a is the best choice.}

@@

>>>>

The default number of bins \(b=b_1\times b_2\) is computed based on the
sample size, by setting \(b_1=n^{1/3}\), consistent with the
Diaconis-Freedman rule \citep{freedman1981}. The value of \(b_2\) is
determined analytically by \(b_1, q, r_2\). Values of \(b_1\) between
\(2\) and \(b_1 = \sqrt{\frac{n}{r_2}}\) are allowed.
Figure~\ref{fig-param-scurve} (b) shows the effect of different choices
of \(b_1\) on the MSE of the fitted model.

@@

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-40-1.pdf}

@@

\caption{long edge removal}

@@

\caption{\label{fig-mnist-mse}Assessing which of the 5 NLDR layouts on
the MNIST digit 1 data is the better representation using MSE for
varying binwidth (\(a_1\)). Colour used for the lines and points in the
left plot and in the scatterplots represents NLDR layout (a-e). All the
layouts appear to be very similar. Layout c is universally poor. Layouts
a that show two close clusters are universally suboptimal. Layout a is
the best choice.}

@@

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-mnist-tri-proj-twisted-1.pdf}

\caption{\label{fig-mnist-tri-proj-twisted}twisted pattern}
